Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence and Forensic Workflow

### 2009-03-11 paper/paper.tex
simson2 Ignore-this: ecbd6ae9f39c5531d3004f8a6bfb7faf
Simson is taking a typing break. This commit is in the middle of a
massive rewrite and refactoring of the paper. Please don't make
substantive changes yet.

% Make sure we know its a draft for now
\usepackage{draftwatermark}
\usepackage{verbatim}
%\documentclass[12pt]{article}
\usepackage{epsfig}
\usepackage[hypertexnames=false,bookmarksopenlevel=1,bookmarksopen,bookmarksnumbered,colorlinks,plainpages=false,linktocpage]{hyperref}
%\usepackage{natbib}
%\pagestyle{empty}
%\psdraft
%\baselineskip=20pt %Sets line spacing to 1 unit
%\bibliographystyle{elsart-num-names}
\bibliographystyle{IEEEtran}
\usepackage{cite}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1:}]}{\end{trivlist}}

@@

This paper extends the previous work by adapting AFF to the multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. We do this in a manner that
is at once upwards compatible with AFF while adopting the AFF
bit-level specification so that AFF files can be processed using
standard non-forensic tools.

\section{A New Forensic Format}

AFF's flexibility comes from the fact that forensic data and metadata
is stored as arbitrary name/value pairs called \emph{segments}. For example, the first 16MB
of a disk image is stored in a segment called \texttt{page0}, the
second 16MB in a segment called \texttt{page1}, \emph{etc.} Because of
this flexibility, it was
relatively easy for Garfinkel to extend AFF to support encryption,
digital signatures, and the storage of new kinds of metadata such as
chain-of-custody information\cite{garfinkel:affcrypto}. .

\subsection{AFF Limitations}
After imaging literally thousands of devices using Garfinkel's AFF
tools and adding support for AFF to numerous open source programs, we
observed numerous problems in the underlying standard
and Garfinkel's AFFLIB implementation:

\begin{itemize}
\item While AFF's design stores a single disk image in each evidence
  file, modern digital investigations typically involve many seized
  computers or pieces of media. 
\item AFF has no provision for storing memory images or intercepted
  network packets.
\item AFF has no provisions for storing extracted files that is
  analogous to the EnCase ``Logical Evidence File'' (L01) format.
\item AFF's encryption system leaks information about the contents of
  an evidence file because segment names are not encrypted.
\item AFF's default compression page size of 16MB can impose significant overhead
  when accessing NTFS Master File Tables (MFT), as these structures
  tend to be highly fragmented on systems that have seen significant
  use. 
\item Although the AFF specification calls for a ``table of contents'' similar
  to the ZIP\cite{zip-format} ``central directory'' that is stored at the end of AFF
  files, Garfinkel never implemented this directory in the publicly
  released AFFLIB. As a result, every header of every segment in an
  AFF file needs to be read when a file is opened. It practice this
  can take 10-30 seconds the first time a large AFF file is
  opened. (Once the file is opened, the sectors corresponding to the
  segment headers are stored in the computer's disk cache.)
\item AFF's bit-level specification is essentially a simple container
  file specification. Given the that there are other container file
  specifications that are much more widely supported with both
  developer and end-user tools, it seemed reasonable to migrate AFF
  from its home-grown format to one of the existing standards.
\end{itemize}

We have addressed these issues with a three step process. First, we
have extended the AFF schema to support multiple forensic images
(which are now called \emph{streams}) in a single evidence
file. Second, we have changed the on disk format, jettisoning AFF1.0's
unique container format and replacing it with the ZIP64 file
format. Third, we are adopting AFFLIB to implement the new schema and
file format, which will allow existing AFF-aware programs to access
files written in the new format without modification. We call this new
format AFF2.

\subsection{AFF2: New Features}
AFF2 is designed to include the following functionality:

\begin{itemize}
\item Evidence is stored in \emph{volumes}. A volume can consist of a
  single file (\emph{e.g.} a file with the \texttt{.aff} extension), multiple files (\texttt{.afd}), an
  object storage system (such as Amazon S3), or a hybrid (for example,
  metadata stored in an SQL database with physical data stored in
  large files in a file system.)
\item A single volume can contain evidence from one or more physical
  devices.
\item RAID information can be directly encoded, allowing data within a
  RAID to be logically constructed at read time.
\item Although the page size remains 16MB by default, each 16MB page
  can be built from multiple compressed chunks (8K size by default),
  significantly decreasing the overhead to access a few blocks inside
  a compressed page.
\item The bit-level container file format has changed from AFFs's
  original container file format to ZIP64, allowing improved
  integration with existing developer tools.
\item Transparent upwards compatibility, so that programs linked with
  the AFFLIB implementation or the pyaff.py python library can
  interchangeably read either AFF1 or AFF2 files. 
\end{itemize}

\section{FIF Specifications}
The FIF specifications distinct between bit level specifications and
a high level schema. The bit level specifications refer to how the
files are to be actually represented on disk. The schema specifies a
stream based abstraction which is implemented using the bit level
specification.

\subsection{FIF bit level specification}
The FIF specifications only require the underlying storage mechanism
be capable of storing multiple named segments of data. The current
specification uses the Zip file format as the underlying storage
mechanism, however in future versions of the specifications other
storage layers may be also be defined.

The Zip file specifications are publicly available and widely
implemented. The Zip64 extensions are provided to enable Zip archives
to grow beyond a two gigabyte limit by using 64 bit
references. Although Zip64 also defines encryption and authentication
extensions, we do not use these in this specification due to the
restrictions imposed on their use. Rather, we specify stream based
encryption scheme for use in FIF files (See section
\ref{crypted_stream}).

The basic structure of a Zip archive is shown in Figure
\ref{zip_structure}. For bit level details see the Zip file
Application notes \cite{zipspecs}.

@@

As can be seen, the archive consists of a {\em Central Directory (CD)}
located at the end of the archive. The CD is a list of pointers to
individual {\em File header} structures located within the body of the
archive. Headers are then followed by the file data, after it has been
compressed by the appropriate compression method (as specified in the
header). Each archived file is optionally followed by a {\em Data
Descriptor} describing the length and CRC of the archived file. Using
the data descriptor field allows implementations to write archives
without needing to seek in the output file. This allows Zip files to
be written to pipes for example, sending an image over the network
using netcat or ssh.

\subsection{FIF Definitions}
The FIF specification defines a number of basic terms:
\begin{itemize}
\item {\em A FIF Volume} is a single ZIP archive containing {\em
segments}.

\item {\em A segment} is a single unit of data written to the Zip
archive as an archive file. The only properties of segments of
interest to the FIF specification are the segment name and the
timestamp.

\item {\em The segment name} is the name under which a segment is
stored within the archive. This consists of a directory part named
{\em a stream} and an {\em chunk name}. For example the segment named
``default/000001'' refers to the stream called ``default'' and has a
chunk name of ``000001''. Although the ZIP specification does not
specify the encoding of segment names, the FIF specification requires
all segment names to be encoded using UTF8. If stream name is
unspecified, implementations should open the stream named ``default''.

\item {\em Attributes} are a list of key/value pairs stored at each
stream's ``properties'' segment (e.g. ``stream/properties''). The
``properties'' file is a text file which stores attributes one on each
line. Attribute lines consist of the key, immediately followed by
``='' for text values or ``:'' for binary values. This is immediately
followed by the encoded value. Values containing text MUST be encoded
using UTF8, while values containing binary data MAY be encoded using
Base64. If keys are repeated, their relative order MUST be preserved
by implementations, but different keys may be presented in any order.

Attributes MUST be specified at the top level of each volume
(i.e. must have a segment named ``properties''). The volume attributes
are used to specify the volume UUID.

%about the stream. A properties file also exists at the top
%level of the ZIP hierarchy to store global information such as the
%volumes UUID, the FIF version and other volumes which make up the

\item {\em A FIF file} is a collection of volumes with the same
UUID. The UUID is generated according to RFC4122 for each new FIF
file. As new volumes are created for the same file, the UUID is copied
to all related volumes. When a FIF file is opened, all its volumes are
loaded and merged in such a way that all the segments from all the
volumes are visible as if they were all found in the same volume.

@@

\label{target_reference}
{\em A target reference} is a way of referencing objects. This is used
in a number of places which will be highlighted below. Using
references it is possible for the FIF file to refer to objects
embedded within itself, or within other FIF files. Its is also
possible to refer to external flat files. A number of usage scenarios
are described in Section \ref{usage_scenarios}. The reference is
specified with the following conventions:

\begin{itemize}
\item A reference starting with {\em file://} refers to an external
file. The external file reference must be sanitised (to remove
.. directory traversal) and rooted at the current directory. A
filename of ``.'' refers to the current FIF volume.

\item A reference in the format {\em fif://UUID/stream/segment} refers
to an external FIF file set (identified by its UUID). If the segment
is missing (i.e. the reference ends with a ``/''), the reference is
for the entire stream. We do not specify a way of actually resolving
the UUID into a physical locations. This can achieved by
implementations using many methods, such as Active Directory, LDAP,
DNS, flat files or SQL databases.

\item A references ending with a ``/'' signify a reference to a
stream in the current FIF file. The stream will be opened in its
entirety. If the reference does not end with a ``/'', the reference is
for a single segment.
\end{itemize}

\subsection{FIF Volumes and segment merging}
There are a number of aspects in the Zip file format specifications
which we do not utilize. For example, the Zip64 specification allows
for a multitude of encryption and hashing algorithms. Some of these
features however, seem to be patent encumbered. Since the FIF
specification aims to separate high level information such as hashes,
cryptography and signatures from the low bit level representation, we
implement those features at a higher level.

In many forensic applications, individual segments within the volume
need to be updated \cite{AFF}. These updates can cause the volume to
become fragmented as holes appear within the middle of the volume, and
updated segments are added at the end.

When we wish to add a new segment to an existing volume, the new
segment header SHOULD overwrite the volume central directory, and the
central directory should be rewritten including the new entry
immediately after the last file in the archive. If it is not possible
to overwrite the file, the new segment can be written at the end,
followed by an updated copy of the central directory (in that case the
old central directory remains in the file). This is illustrated in
Figure \ref{zip_structure}. Implementations may choose to delay
writing the central directory until the file is closed.

A segment within the archive is considered to be updated when a new
segment is added to the archive with the same name and a later
timestamp. Implementations MUST satisfy read requests for duplicate
segment names with those segments with the later timestamp, or if the
two timestamps are the same, the segment which appears at a later
offset in the volume. This effectively allows for a segment to be
updated without leading to internal volume fragmentation.

The fact that the older version of all updated segments remains intact
within each volume can be used for auditing, or to allow a rollback to
previous versions of the segment. The {\em frepack} tool is provided
to repack, split and manage FIF volumes by creating new volumes and
redistributing segments among them.

Note that we do not rely on Zip64's built-in support for splitting
archives at all (although implementations should be able to read such
files). Each volume is a complete and stand alone ZIP file. It is the
FIF implementation which considers the segments contained within it to
belong to a larger collection. The reason behind this decision is that
often it is needed to add new volumes to an existing set without
modifying the existing set at all.

Consider the typical usage scenario depicted in Figure \ref{usage}, of
a FIF file which contains a disk image. This volume is distributed to
two independent analysts, which wish to add new streams to the FIF
file (for example the result of their analysis). In this case they can
each create a new volume which extends to original volume and save
their analysis on this new volume. Now they only need to share this
new volume with other analysts who also have a copy of the old volume
to interchange their findings.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{usage1.eps}}

  \caption{A typical usage scenario. Both Alice and Bob receive a FIF
  volume but work independently. Rather than modifying the volume,
  they each create another volume and save their results to the new
  volume. They can now exchange the smaller new volumes and
  effectively merge their results into the same FIF set.}

  \label{usage}
  \end{center}
\end{figure}

This is made possible because each volume is independent of one
another, and can be loosely collated into a bigger FIF file
set.

\subsection{FIF Streams}
The above section explained how FIF file sets can be split into
volumes, each containing segments. When the entire FIF file set is
loaded, segments may be fetched from any volume. The present section
describes how these individual segments are used to form {\em stream
abstractions}.

A stream is a high level abstraction of a single source of data. For
example, a stream may contain a disk image or a network
capture. Although we do not specify the exact API, implementations
MUST make a stream opened for reading have at a minimum a read, seek
and tell methods. Seeking MUST support the whence argument as
described in the standard C implementation (i.e. allow seeking to the
end of the stream).

Streams opened for writing MAY limit support for seeking, as efficient
seeking in compressed streams may not be possible. Applications should
avoid needing to seek in output streams, and try to structure writes
serially. It is usually possible to reorder a sequence of non
overlapping random writes into an Image stream and a Map Stream (These
are discussed in Sections \ref{image_stream} and
\ref{map_stream}). Implementations are free to hide the inability to
seek in writing streams by automatically implementing such a scheme,
but this is not mandated by the specification.

All streams within the FIF file MUST have a properties segment. The
properties segment describes the stream attributes as specified
previously. Streams must have the {\em type} attribute which allows
implementations to determine how they should be processed.

\section{Standard FIF segments}
The FIF specification defines a number of basic types, but extensions
are encouraged. Future versions of the specification might standardize
further types. Implementations MUST implement the basic types.

\subsection{The Image Stream}
\label{image_stream}
The {\em Image} stream ({\em type=image}) contains a single read only
forensic data source. For example, this Stream might contains a hard
disk image, a memory image or a network capture (in PCAP format). The
Image stream provides random read access into the stream (i.e. allows
for randomly seeking the stream to any point). The {\em size}
attribute MUST be specified to determine how large the stream is.

The image itself is stored in chunks within the FIF file. Chunks are
segments named as an 8 digit, zero padded decimal integer
representation of their chunk id (e.g. ``default/00000032''). Image
streams must specify the {\em chunk\_size} attribute, as the number of
image bytes each chunk contains. Chunks may be compressed at the
underlying Zip container level, using any of the compression
algorithms specified by the zip specifications.

Chunks can optionally be individually hashed, allowing for detection
of modifications. The attribute {\em hash\_type} specifies the type of
hash used (implementations SHOULD support md5,sha1 and sha256, but can
define others). The hashes are stored in any number of segments within
the stream, of arbitrary name. The hash segment names is provided by
using the {\em hash} stream attribute.

Each hash segment MUST be a list of a chunk id encoded in 32 bit
little endian integer format followed by the hash (the length of which
depends on the hash itself). This approach is illustrated in Figure
\ref{hash_index}.

Hash segments can contain any number of chunk hashes in them. It is
recommended that a hash segment be created when a volume is closed to
cover all the chunks within that volume. This allows for hashes to be
verified even if some volumes are missing. It is permitted for a hash
segment to update a hash of a chunk previously set by an earlier hash
segment - this may be required when a chunk is updated. The hash
segment can then be signed to assure the integrity of the updated
segment.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}

  \caption{The structure of a hashing segment. The attributes {\em
  hash} and {\em hash\_type} declare the hash segment. The segment
  contains a list of integer encoded chunk id followed by an unencoded
  hash for each chunk.}

  \label{hash_index}
  \end{center}
\end{figure}

This design differs from AFF, which creates a separate segment for
each hash. The Zip headers are more verbose than the AFF segment
headers, and typically FIF chunks are much smaller (default 32kb, vs
16Mb for AFF). Therefore, creating a new segment for each hash value
results in more overheads in the FIF format. By packing the hashes
into a few segments it is possible to reduce this overhead to a
minimum.

Hashes are calculated in an identical way to that defined by AFF,
thats is the full name of the segment (including stream component) is
followed by a single NULL byte, followed by the (uncompressed) content
of the chunk. The inclusion of the segment name ensure that even
identical chunks in content will have a different hash.

@@

>>>>

% Make sure we know its a draft for now
\usepackage{draftwatermark}
\SetWatermarkFontSize{3cm}
\usepackage{verbatim}
%\documentclass[12pt]{article}
\usepackage{epsfig}
\usepackage[hypertexnames=false,bookmarksopenlevel=1,bookmarksopen,bookmarksnumbered,colorlinks,plainpages=false,linktocpage,linkcolor=black,citecolor=black,filecolor=black,urlcolor=black]{hyperref}
%\usepackage{natbib}
%\pagestyle{empty}
%\psdraft
%\baselineskip=20pt %Sets line spacing to 1 unit
%\bibliographystyle{elsart-num-names}
\bibliographystyle{IEEEtran}
\usepackage{cite}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1:}]}{\end{trivlist}}

@@

This paper extends the previous work by adapting AFF to the multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. We do this in a manner that
is at once upwards compatible with AFF while adopting the AFF
bit-level specification so that AFF files can be processed using
standard non-forensic tools.  We call the new system AFF4, and use the
phrase AFF1 to refer to the legacy system developed by
Garfinkel.\footnote{Although Garfinkel never changed the AFF bit-level
specification, Garfinkel released AFFLIB implementations with major
version numbers 1, 2 and 3. We therefore call our system AFF4 to avoid confusion.}

\begin{figure}
\fbox{
\begin{minipage}{.95\linewidth}
\begin{itemize}
\item The ability to store evidence from one or more physical
  devices.
\item Schema allowing evidence files to contain memory
  images, network packet intercepts, Netflow data, and extracted
  files.
\item Direct encoding of RAID configuration, allowing RAID volumes
  to be automatically reassembled from evidence files made of the
  physical RAID devices.
\item A revised compression system that allows individual chunks
  within a 16MB segment to be decompressed without decompressing the
  entire segment. 
\item Support for Global Distributed Evidence Management.
\end{itemize}
\end{minipage}}
\caption{Specific design goals for the next generation forensic file
  format.}
\end{figure}
\section{The Need for an Improved Forensic Format}

AFF1's flexibility came from the fact that forensic data and metadata
is stored as arbitrary name/value pairs called \emph{segments}. For
example, the first 16MB of a disk image is stored in a segment called
\texttt{page0}, the second 16MB in a segment called \texttt{page1},
\emph{etc.} Because of this flexibility, it was relatively easy for
Garfinkel to extend AFF1 to support encryption, digital signatures, and
the storage of new kinds of metadata such as chain-of-custody
information\cite{garfinkel:affcrypto}.

\subsection{AFF Limitations}
After imaging literally thousands of devices using Garfinkel's AFF1
tools and adding support for AFF1 to numerous open source programs, we
observed numerous problems in the underlying standard
and Garfinkel's AFFLIB implementation:

\begin{itemize}
\item While AFF1's design stores a single disk image in each evidence
  file, modern digital investigations typically involve many seized
  computers or pieces of media. 
\item AFF1 has no provision for storing memory images or intercepted
  network packets.
\item AFF1 has no provisions for storing extracted files that is
  analogous to the EnCase ``Logical Evidence File'' (L01) format, or
  for linking evidence to web pages.
\item AFF1's encryption system leaks information about the contents of
  an evidence file because segment names are not encrypted.
\item AFF1's default compression page size of 16MB can impose significant overhead
  when accessing NTFS Master File Tables (MFT), as these structures
  tend to be highly fragmented on systems that have seen significant
  use. 
\item Although the AFF1 specification calls for a ``table of contents'' similar
  to the ZIP\cite{zip-format} ``central directory'' that is stored at the end of AFF
  files, Garfinkel never implemented this directory in the publicly
  released AFF1 implementation, AFFLIB. As a result, every header of every segment in an
  AFF file needs to be read when a file is opened. It practice this
  can take 10--30 seconds the first time a large AFF file is
  opened. (Once the file is opened, the sectors corresponding to the
  segment headers are stored in the computer's disk cache.)
\item AFF1's bit-level specification is essentially a simple container
  file specification. Given the that there are other container file
  specifications that are much more widely supported with both
  developer and end-user tools, it seemed reasonable to migrate AFF
  from its home-grown format to one of the existing standards. 
\end{itemize}

\subsection{Global Distributed Evidence Management}
While Garfinkel designed AFF for use on a single machine that could
both image evidence and perform analysis, the main use of AFF in
practice has been in distributed environments in which imaging and
analysis takes place in multiple locations and is performed by
multiple individuals.

Managing evidence in a globally distributed system requires the use of
globally unique identifiers. The original AFF specification assigned
each piece of evidence a unique 128-bit identifier called a GID but did
not make it clear when this identifier should be changed and when it
should remain the same.

In practice, however, global distributed evidence management requires
more than simply tracking the movement of disk images: it requires
approaches for sharing evidence to multiple disconnected
evidence, allowing offline work, and then seamlessly recombining the
work products of the analysts in a third security domain.

Consider the typical usage scenario depicted in Figure \ref{usage}, of
an file which contains a disk image. This volume is distributed to two
independent analysts, Alice and Bob. Alice may find and extract
individual files, while Bob may correlate information in the evidence
file with other data that is available on departmental
servers. Although in some environments Alice and Bob may be able to
work on a shared file that is located on a server, in other
environments there will not be sufficient bandwidth or
connectivity. Instead, each analyst will be required to store the
information in their own evidence file; these files will then be
recombined at a later point in time.

In this case they can
each create a new volume which extends to original volume and save
their analysis on this new volume. Now they only need to share this
new volume with other analysts who also have a copy of the old volume
to interchange their findings.

This is made possible because each volume is independent of one
another, but can still be recombined into a bigger evidence set.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{usage1.eps}}
  \caption{A typical usage scenario. Both Alice and Bob receive an AFF
  volume but work independently. Rather than modifying the volume,
  they each create their own local volumes and save their results into
  those files. They can now exchange the smaller new volumes and
  effectively merge their results into the same AFF set when they are finished.}
  \label{usage}
  \end{center}
\end{figure}

\subsection{Migration from AFF1 to AFF4}

We have addressed these issues in three steps. First, we have extended
the AFF schema to support multiple forensic images (which are now
called \emph{streams}) in a single evidence file. Second, we replaced
AFF's ``bit-level'' format with an improved format that is more
flexible and has better existing support. Third, we are adopting
AFFLIB to implement the new schema and file format, which will allow
existing AFF-aware programs to access files written in either the old
or the new format without modification.

We call the result of our work effort AFF4.

\section{Introducing AFF4}

\subsection{AFF4 Terminology and New Concepts}

This section discusses the AFF4 terminology.

\begin{itemize}
\item \emph{An Evidence Volume} is a single file with a \texttt{.aff}
  extension. 
\item \emph{An Evidence Set} is a collection of one or more Evidence
  Volumes. Evidence Sets can also be stored in object storage system (such as
  Amazon S3), or a hybrid---for example, metadata stored in an SQL
  database with physical data stored in large files in a file
  system. 
\item \emph{An Evidence Set UUID} is a unique identifier generated
  according to RFC4122\cite{rfc4122} for each new Evidence
  Set. Multiple Evidence Volumes that make up the same Evidence Set
  should all have the same UUID. When an Evidence Set is opened all its volumes are
  loaded, their UUIDs are checked, and all of their contents are
  logically merged so as to be visible as if they were all found in the same volume.
\item \emph{A segment} is a single unit of data written to volume. AFF4
  segments have a \emph{segment name}, a \emph{segment timestamp}, and
  a \emph{segment contents}.

\item \emph{A stream} is a collection of segments that can be
  concatenated together to form a single data object.

\item \emph{An image stream} is a stream that is dedicated to storing evidence from a single
  source of data.  For example, a stream may contain a disk image or a network
  capture. Whereas AFF could only support a single stream in an evidence
  file, AFF4 allows any number of streams to be stored. Each stream in
  the archive is  given a unique identifier.

\item \emph{A bevy} is a specific segment that is used to hold
  forensic image data, such as disk sectors, memory pages, or network
  packets. AFF1 used the term ``pages'' to describe bevys, but this
  term caused confusion when AFF was used to hold memory images, hence the
  new term.

\item \emph{A segment name} is the name under which a segment is
  stored within the archive. AFF4 segments have two-part names that
  typically contain a \emph{stream name} and \emph{bevy number}. For
  example the segment named \texttt{default/000001} refers to the stream
  called \texttt{default} and has a chunk name of \texttt{000001}. All segment
  names are encoded using UTF8. AFFLIB is being modified to allow any
  stream to be opened using the \texttt{af\_open} function; if a
  stream name is not specified, AFF4 implementations will open the
  stream named ``default.''

\item \emph{The segment timestamp} is a timestamp for the segment
  recorded in GMT.

\item \emph{Attributes} are a list of key/value pairs stored in each
stream's ``properties'' segment (e.g. ``stream/properties''). The
``properties'' file is a text file which stores attributes one on each
line. Attribute lines consist of the key, immediately followed by
``='' for text values or ``:'' for binary values. This is immediately
followed by the encoded value. Values containing text are encoded
using UTF8, while values containing binary data are encoded using
Base64. There may be multiple values for each attribute; in these
cases the order is preserved
by implementations. Implements may present different keys may be
presented in any order, however.
\vspace{1ex}
Each AFF4 archive must also have a top-level set of attributes---that
is, there is always a segment named ``properties.'' These volume
attributes are used to specify a Unique Identifier (UUID). (AFF1
stored each property in its own named segment. This created
significant overhead in many situations.)

\item \emph{A Target Reference} is a way of referencing objects. This is used
in a number of places which will be highlighted below. Using
references it is possible for the FIF file to refer to objects
embedded within itself, or within other FIF files. Its is also
possible to refer to external flat files. A number of usage scenarios
are described in Section \ref{usage_scenarios}. 
\end{itemize}

The Target Reference is an important AFF4 innovation: this mechanism
gives AFF4 files a way for describing evidence that is stored in
different evidence volumes, in the host computer's file system, or on
the Internet. We have defined the following target reference schema:

\label{target_reference}

\begin{itemize}
\item A reference in the format {\em aff://UUID/stream/segment} refers
to an external Evidence file set (identified by its UUID). If the segment
is missing (i.e. the reference ends with a ``/''), the reference is
for the entire stream. We do not specify a way of actually resolving
the UUID into a physical locations. This can achieved by
implementations using many methods, such as Active Directory, LDAP,
DNS, flat files or SQL databases.

\item A reference starting with {\em file://} refers to an external
file. The external file reference must be sanitised (to remove
.. directory traversal) and rooted at the current directory. A
filename of ``.'' refers to the current FIF volume.

\item A references ending with a ``/'' signifies a reference to a
stream in the current FIF file. The stream will be opened in its
entirety. If the reference does not end with a ``/'', the reference is
for a single segment.
\end{itemize}

\subsection{AFF4 Streams}
As discussed above, individual segments in an AFF4 evidence set can be
combined into a single logical object called a \emph{stream}. AFF4
supports an arbitrary number of streams per volume. The AFFLIB library
allows the segments in a stream to be operated on as if they were a
single file inside a file system by supporting the traditional
POSIX-like functionality of \texttt{open()}, \texttt{seek()} and
\texttt{read()}.

AFFLIB also supports the \emph{writing} of streams, a feature that is
used by Garfinkel's \texttt{aimage} disk imaging program. In general
arbitrary writing to streams can be expensive, especially when writing
is combined with seek operations. The current \texttt{aimage} program
will create AFF files with gaps in them during its error recovery
process (when \texttt{aimage} seeks to the end of a disk and reads it
backwards in an attempt to get around media failures). We are working
on a rewrite of \texttt{aimage} which performs imaging and recovery
on bevy boundaries, rather than using the current arbitrary
``high-water-mark'' and ``low-water-mark.''

All streams within an Evidence Set \emph{must} have a properties segment. The
properties segment describes what the stream contains and determines
how the stream is processed.

\subsection{The Image Stream}
\label{image_stream}
The AFF4 \emph{Image} stream contains a single read-only forensic data
set. For example, this stream might contains a hard
disk image, a memory image or a network capture (in PCAP
format). Random read access into the stream is provided with
AFFLIB. (AFFLIB also provided the ability to randomly \emph{write}
into the evidence stream for sanitization purposes, although this
feature may be removed at a later date.)

Image streams must have at several attributes. First, they must have
\texttt{type=image} to indicate that they are image streams. Streams
must have a \texttt{size} attribute (specified in 8-bit bytes) to
indicate the last addressable byte in the stream. Streams can have
holes in them; in this case, attempts to read data from within the
holes can generate an error or return NULLs.

The image itself is stored in \emph{pages} within the AFF4 file. Pages are
segments named as an 8 digit, zero padded decimal integer
representation of their chunk id (e.g. ``default/00000032''). Image
streams must specify the \texttt{page\_size} attribute, as the number of
image bytes each chunk contains. Pages may be compressed at the
page level, or may be broken up into multiple chunks (typically 8KB),
and individually compressed. In this case the beginning of the page
segment will contain an offset table indicating the location of each
compression chunk. This is similar to the technique used by the Expert
Witness file format used by EnCase\cite{encase-3.0} and implemented by
the open source libewf\cite{libewf} package.

Pages can optionally be individually hashed, allowing for detection
of modifications. The attribute \texttt{hash\_type} specifies the type of
hash used (implementations will typically support md5,sha1 and sha256, but can
define others). The hashes are stored in any number of segments within
the stream, of arbitrary name. The hash segment names is provided by
using the \texttt{hash\_name} stream attribute.

Hashes are calculated in an identical way to that defined by AFF1's
crypto layer\cite{garfinkel:affcrypto}---that is, the full name of the
segment (including stream component) is followed by a single NULL
byte, followed by the (uncompressed) content of the chunk. The
inclusion of the segment name ensure that even identical chunks in
content will have a different hash and prevents reordering attacks.

Each hash segment is a list of a chunk id encoded in 32 bit
little endian integer format followed by the hash (the length of which
depends on the hash itself). This approach is illustrated in Figure
\ref{hash_index}.

Hash segments can contain any number of chunk hashes in them. It is
recommended that a hash segment be created when a volume is closed to
cover all the chunks within that volume. This allows for hashes to be
verified even if some volumes are missing. The AFFLIB implementation
may update a hash of a chunk previously set by an earlier hash
segment---this would be required when a chunk is updated. The hash
segment can then be signed to assure the integrity of the updated segment.

\begin{figure}[tb]
  \begin{center}
  \mbox{\epsfxsize=0.6\columnwidth \epsffile{hash.eps}}
  \caption{The structure of a hashing segment. The attributes {\em
  hash} and {\em hash\_type} declare the hash segment. The segment
  contains a list of integer encoded chunk id followed by an unencoded
  hash for each chunk.}
  \label{hash_index}
  \end{center}
\end{figure}

This design differs from AFF1, which creates a separate segment for
each hash. AFF4 uses headers that are more verbose than the AFF
segment headers. Therefore, creating a new segment for each hash value
would result in too much overhead: Packing the hashes into a few
segments it is possible to reduce this overhead to a minimum.

\subsection{ZIP64}
For AFF4, we have decided to change AFF1's
original container file format to ZIP64\cite{zipspecs}. There are many
reasons for this decision:

\begin{itemize}
\item There is already wide support for the ZIP and ZIP64 formats. By
  migrating to these formats, we can take advantage of the rich number
  of user and developer tools already available.
\item ZIP64 includes a CRC32 and timestamp for each segment. We
  believe that the additional integrity checks and timestamps will be
  useful in a forensic environment.
\item ZIP64 already includes support for the ``table of contents''
  envisioned but never implemented by Garfinkel; ZIP64 calls this
  table the ``Central Directory.'' It is written at the end of the
  ZIP64 volume.
\end{itemize}

Figure~\ref{zip_structure} shows the basic structure of a Zip
archive. As can be seen, the archive consists of a {\em Central Directory} (CD)
located at the end of the archive. The CD is a list of pointers to
individual {\em File header} structures located within the body of the
archive. Headers are then followed by the file data, after it has been
compressed by the appropriate compression method (as specified in the
header). Each archived file is optionally followed by a {\em Data
Descriptor} describing the length and CRC of the archived file. Using
the data descriptor field allows implementations to write archives
without needing to seek in the output file. This allows Zip files to
be written to pipes for example, sending an image over the network
using netcat or ssh.

@@

It is important to note that AFF4 only requires that the underlying storage mechanism
be capable of storing multiple named segments of data. Although our
AFF4 implementation will use the ZIP64 file format as an underlying storage
mechanism, our system also supports legacy AFF1 volumes as well as
segments that are stored in other storage systems---for example,
Amazon.com's Simple Storage System (S3)\cite{s2-aws-home-page-money}.

\subsubsection{Unused ZIP64 Features}

There are a number of aspects in the Zip file format specifications
which we do not utilize. 
\begin{itemize}
\item We ignore Zip64's built-in support for splitting
archives into multiple ZIP files. Instead, our implementation treats
each volume as complete and stand-alone ZIP file. The AFF4
implementation then considers the segments contained within as 
belonging to a larger collection. The reason behind this decision is that
often it is needed to add new volumes to an existing set without
modifying the existing set at all. Zip64's built-in support for
splitting an archive renders each individual volume in the set
unusable unless all are available.

\item Although Zip64 also defines encryption and authentication
extensions, we do not use them due to the
restrictions imposed on their use and because they lack of functionality that is
important for forensic user. Instead, we use AFF1's digital signature
facilities for integrity and non-repudiation, and we introduce a new  stream based
encryption scheme for ensuring data privacy (Section~\ref{crypted_stream}).

@@

\subsubsection{A Forensicly Sound Strategy For Updating ZIP64 Files}

In many forensic applications, individual segments within the volume
need to be updated. To this end, Garfinkel implemented several
``update'' functions in AFF1\cite{garfinkel:aff}.

There were two problems with the original AFF updating approach:
\begin{itemize}
\item Updates can cause the volume to
become fragmented as holes appear within the middle of the volume, and
updated segments are added at the end.
\item Updates could be performed silently, raising questions of
  forensic soundness.
\end{itemize}

In AFF4, when a new segment is added to an existing volume, the new
segment header overwrites the volume central directory, and the
central directory should be rewritten including the new entry
immediately after the last file in the archive. If it is not possible
to overwrite the file, the new segment can be written at the end,
followed by an updated copy of the central directory (in that case the
old central directory remains in the file). This is illustrated in
Figure \ref{zip_structure}. Implementations may delay
writing the central directory until the file is closed; should the
system crash before the central directory is written, it can be
recovered by scanning the entire ZIP file.

A segment within the archive is considered to be updated when a new
segment is added to the archive with the same name and a later
timestamp. Implementations satisfy read requests for duplicate
segment names with those segments with the later timestamp, or if the
two timestamps are the same, the segment which appears at a later
offset in the volume. This effectively allows for a segment to be
updated without leading to internal volume fragmentation.

The fact that the older version of all updated segments remains intact
within each volume can be used for auditing, or to allow a rollback to
previous versions of the segment. We have created a new tool,
\texttt{afrepack}, which can repack, split and manage AFF4 volumes by
creating new volumes and redistributing segments among them.

\subsubsection{Segment merging}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Standard FIF segments}
The FIF specification defines a number of basic types, but extensions
are encouraged. Future versions of the specification might standardize
further types. Implementations MUST implement the basic types.

@@

\section{Conclusion and Future Work}
\subsection{Expanding AFFLIB for AFF4}
Although we do not specify the exact API, implementations
MUST make a stream opened for reading have at a minimum a read, seek
and tell methods. Seeking MUST support the whence argument as
described in the standard C implementation (i.e. allow seeking to the
end of the stream).
